### 2018-05-15 data_sets.tex

\subsection{Graph Data Sets}

Before we focus on graph data generators, we briefly overview graph data sets devoted to benchmarking of various types of graph algorithms. \cite{Knuth:1993:SGP:164984} denotes such a  database of graphs (or graph library or graph archive) and related computer programs as a \emph{graphbase}. In general, there exist several types of testing graph data sets~\cite{Bachmaier2012}:

\begin{itemize}
  \item A real-world data set which represents a graph, such as the Internet Movie Database (IMDb)\footnote{\url{http://www.imdb.com/}} or various data sets available in the Stanford Large Network Dataset Collection~\cite{snapnets},
  \item A general graphbase, such as the Open Graph Archive~\cite{Bachmaier2012}, the Stanford GraphBase~\cite{Knuth:1993:SGP:164984}, or the GraphArchive~\cite{GraphArchive},
  \item A dataset dedicated to specific experimental goals, such as graph partitioning and graph clustering~\cite{10dimacs}, shortest paths~\cite{9dimacs}, or graph coloring~\cite{coloring}, and
  \item A dataset devoted to matrices, such as the SuiteSparse Matrix Collection~\cite{SuiteSparse} or  the Matrix Market~\cite{MatrixMarket}.
\end{itemize}

...