### 2018-05-15 generators.tex

\section{Graph Data Generators}
\label{sec:generators}

...

\input{data_sets}

\input{generators_general}

\input{generators_LinkedData}

\input{generators_GraphDatabases}

\input{generators_SocialNetworks}

\input{generators_analytics}

\input{generators_streaming}

% \subsection{Specific Types of Graphs}
% ...

% % Note: not in timeline or tables
% \paragraph{Heterogeneous Graphs} The Heterogeneous Graph Data Benchmark (GDB-H)~\cite{Gupta:2012:GLH:2741795.2741808} aims at \emph{heterogenous graph} data model, a mixed model graph structure that combines several existing generation techniques into a single benchmark. The idea is demonstrated using a drug discovery scenario whose schema involves 11 entity categories (e.g., genes, proteins, diseases, ...) and 3000 binary relationships (e.g., instanceOf, subclassOf, ...). The data is structured as a combination of $N$ overlapping named graphs $G_1, ... G_N$, where the overlap is accomplished by node sharing. A subset of the named graphs $G_1, ... G_k$ are hierarchical, i.e., they are structured as trees or DAGs. The remaining $N-k$ graphs are multigraphs which differ in terms of their network connectivity properties (some component graphs obey the power-law more strictly, some graphs have a larger skew in the distribution of edge labels, some graphs  are denser, some graphs may optionally have additional constraints regarding subgraph patterns). The user can specify the number of component graphs (8 to 100), the number of nodes (100,000 to 100,000,000), the number of node types (3 to 11), and the number of distinct edge labels (30 - 3000), and optionally also type ratios (the fraction of the component graphs having hierarchical, power-law, community or motif structure), node distribution (i.e. the relative size of graph components), edge density, overlapping etc.

% For the purpose of generating the heterogeneous graphs having heterogeneous, i.e. hierarchical, power-law, community-structured or purely random, structure the authors combine several existing approaches corresponding to the particular structural type~\cite{PhysRevLett.102.128701,doi:10.1080/10586458.2001.10504428}.

% \subsection{Object-Oriented and XML Databases (?)}

% see~\cite{Dominguez-Sal:2010:DDG:1946050.1946053}