### 2018-07-17 challenges.tex
Added some braindump about open challenges

\section{Challenges and Open Problems}
\label{sec:challenges}

In this section, we discuss several of the open challenges in large scale graph
generation.

\subsection{Single- vs. multi-domain}

Most of existing graph generators generate graphs that are either not labeled or
are specific of a given domain (e.g.  Social Networks). Graphs from different
domains have different schemas, structural characteristics, property
distributions, etc. which might have an impact on the performance of the
application under test. Thus, graph processing engine developers are asking for
generators or tools to generate graphs flexibily, allowing to configure aspects
such as the size, the schema, the data distributions and other structural
characteristics such as the degree distribution, the clustering coefficient,
etc.

\subsection{Simple Usage, Simple Parameters}
Proposal of a data generator (not necessarily for graph data) has to face an
important schism. On the one hand, there is the aim to provide the user with as
many parameters as possible in order to enable one to generate any kind of data.
This approach seems to be reasonable, but it encounters the fact that users are
usually unwilling to use complex benchmarking tools. This observation can be
seen, for example, in the case of XML benchmarks -- even though there exist
robust and complex data generators (such as
ToXGene~\cite{conf/webdb/BarbosaMKL02}, which supports specification of
structural aspects, value distributions, references etc.), the most popular
benchmarking tool is XMark~\cite{Schmidt:2002:XBX:1287369.1287455} which models
a single use case and enables to specify just the size of the data. Hence, the
other extreme is to provide a simple data generator which does not require any
complex settings and thus the benchmarking process is extremely simple and fast.

Considering the complex structure of graph data and the variety of applications
having highly specific types of graphs, the latter solution is difficult to
implement. A reasonable compromise can thus be seen in a data generator which is
provided with sample graph data and is capable of automatic analysis of their
structural and value specifics in order to gain the complex parameters.

\subsection{Evolving Graph Data}
As user requirements as well as environments change, most of the existing
applications naturally evolve over time, at least to some extent. This evolution
usually influences the structure of the data and consequently all the related
parts of the application (i.e., storage strategies, operations, indices etc.).
The respective data generator should hence be able to simulate such a growth
and/or change in structure.

In some graph applications, such as, e.g., social networks, the evolution of the
data is a significant aspect, especially in the activity graphs, that has been
studied
extensively~\cite{doreian1997evolution,Kumar:2006:SEO:1150402.1150476,Hellmann2014583,wang2013,Kossinets88,Viswanath:2009:EUI:1592665.1592675}.
A related problem is \emph{data versioning} and respective ability to query
across multiple versions of data or in general its analysis. Considering graph
data this problem is also common, for example  in the area of Linked
Data~\cite{DBLP:conf/semweb/Papakonstantinou16,DBLP:conf/esws/MeimarisP16,fernandez2015towards,fernandez2015bear}.

As shown in
papers~\cite{Leskovec:2005:RMT:2101235.2101254,Leskovec:2005:GOT:1081870.1081893},
evolving graphs have further specific features. For example, some graphs grow
over time according to a \emph{densification power law} which means that real
graphs tend to sprout many more edges than nodes, and thus are densifying as
they grow. Also the effective diameter of graphs tends to shrink or stabilize as
the graph grows with time.

%\paragraph{} Following the same recursion idea,
%paper~\cite{Leskovec:2005:RMT:2101235.2101254} uses Kronecker multiplication to
%generate self-similar graphs. The network starts with an initial graph G1 that
%contains $N_1$ nodes and $E_1$ edges. Using matrix recursion, larger successive
%graphs $G_2, G_3, ... G_n$ are generated. The $k$-th graph $G_k$ contains $N_k
%= N^k_1$ nodes. Many graphs often densify over time, exhibiting a growth in the
%number of edges that is superlinear to the number of nodes. Kronecker
%multiplication produces graphs with a fixed diameter and a densification power
%law degree distribution with exponent $k = log(E_1)/log(N_1)$. The graph
%generation process introduces a staircase effect in the nodes' degrees, and
%each community consists of smaller nested communities that are formed through
%expansion and recursion.

\subsection{Multi-model Data}
With the dawn of Big Data and especially its Variety aspect there have emerged
also new types of database management systems. One of the most interesting ones
are so-called \emph{multi-model databases} which enable to store and especially
query across structurally different data. There exist various types of
multi-model systems combining  distinct subsets of Big Data structures; hence,
there naturally exists systems which combine graph data with other data models.
For example, OrientDB\footnote{\url{http://orientdb.com/orientdb/}} which was
implemented on the basis of an object DBMS currently supports graph, document,
key/value, and object models.

Such type of DBMSs also needs a specific benchmark/data generator which enables
to test new features and analyze efficiency of operations. However, since the
multi-model systems are in the context of Big Data rather new, there exist only
a few benchmarks targeting multi-model DBMSs (such as
Bigframe~\cite{journals/pvldb/KunjirKB14} or UniBench~\cite{conf/cidr/lu17})
with limited capabilities.

\subsection{Large scale graphs with realistic structure}

Most of existing graph generators aiming at generating large graphs with realistic
structural characteristics focus principally on reproducing the degree
distribution and the clustering
coefficient~\cite{kolda2014scalable,edunov2016darwini}. However, there are other
structural characteristics one might be interested in reproducing for a large
graph, such the diameter, the size of the largest connected component, or the
hierarchical community structure. Graph practitioners are intersted in knowing
how other high-level structural characteristics affect the performance of graph
queries and graph algorithms. Thus, there is an open challenge to create
graph generators that allow reproducing many other structural characteristics
besides the degree or the clustering coefficient, that can also generat graphs
at a scale.

\subsection{Machine Learning based graph generation}

With the advent of neural networks and specially generative adversarial networks
(GANs)~\cite{goodfellow2014generative}, several researchers have started to
explore their application to generate graphs. This is the case of
~\cite{kipf2016variational,grover2018graphite,simonovsky2018graphvae,li2018learning,you2018graphrnn},
which present several generative models to generate realistic graphs.  Such
techniques still suffer from several problems. For instance, some of them are
limited to learn from a single
graph~\cite{kipf2016variational,grover2018graphite} or generate small
graphs~\cite{simonovsky2018graphvae,li2018learning,you2018graphrnn}. The
technique proposed in~\cite{you2018graphrnn} is capable of generating graphs
with complex edge dependencies (e.g. community structure) and is not restricted
to graphs of a fixed size. However, there is still in general several open
challenges, including the capability of learning from and generating large graphs comparable
in size to those typically used for benchmarking, and robust generation
techniques with structural guarantees (e.g. degree distribution, clustering
coefficient, etc.).