### 2018-09-24 generators_community_detection.tex

\subsubsection{Community Detection}
\label{sec:generators_community_detection}

Community detection is one of the many graph analytics algorithms typically used on domains such as social networks or bioinformatics. Communities are groups of vertices that are highly connected among them, while being scarcely connected to the rest of the graph. Such communities emerge from the fact that real graphs are not random, but follow real-world dynamics that make similar entities to have a larger probability to be connected. As a consequence, detected communities are used to reveal vertices with similar characteristics, for instance to discover functionally equivalent proteins in protein-protein interaction networks, or persons with similar interests in social networks. Such applications have made community detection a hot topic during the last fifteen years with tens of developed algorithms and detection strategies~\cite{doi:10.1002/wics.1403,Kim:2015:CDM:2854006.2854013}. In order to compare the quality of the different proposed techniques, one needs graphs with \emph{reference communities}, that is, communities known beforehand. Since it is very difficult to have large real graphs with reference communities (mainly because these would require a manual labeling), graphs for benchmarking community detection algorithms are typically generated synthetically.

The first attempts to compare community detection algorithms using synthetic graphs proposed the use of random graphs composed by several Erd\"{o}s-R\'{e}nyi subgraphs, connected more internally than externally~\cite{danon2005comparing}. Each of these subgraphs has the same size and the same internal/external density of edges. However, such graphs miss the realism observed in real graphs, where communities are of different sizes and densities. Thus, Lancichinetti, Fortunato and Radicchi (hence LFR)~\cite{PhysRevE.78.046110} propose a class of benchmark graphs for community detection where communities are of different sizes and densities. The generator generates communities of different sizes following a power-law distribution whose parameters can be configured. The degree of the nodes is also sampled from a power-law distribution. Additionally, the generator introduces the concept of the ``mixing factor'', which specifies the fraction of its links connecting nodes belonging to different communities. Such parameter allows  the degree of modularity of the generated graph  to be tuned, thus testing the robustness of the algorithms under different conditions. The generation process starts with an empty graph and incrementally fills in the adjacency matrix by obeying the described constraints. Lancichinetti, Fortunato and Radicchi~\cite{PhysRevE.80.016118} extended LFR to support the notion of directed graphs and overlapping communities. Overlapping communities extend the notion of communities by allowing the sharing of vertices, thus a vertex can belong to more than one community. This extended generator allows controlling the same parameters of the previous version, but also the amount of overlap of the generated communities.

Besides synthetic graph generators, Yang and Leskovec~\cite{yang2015defining} proposed the use of real-world graphs with explicit group annotations (e.g., forums in a social network, categories of products, etc.) to infer what they call ``meta-communities'', and use them to evaluate overlapping community detection algorithms. However, a recent study from Hric, Darst and Fortunato~\cite{hric2014community} reveal a loose correspondence between  communities (the authors refer to them as structural communities) and meta-communities.  This result reveals that  algorithms working for structural communities do not work well for finding meta-communities and vice versa, suggesting significantly different underlying characteristics between the two types of communities, which are yet to be identified.

In this regard and to the best of our knowledge, there does not exist a set of generators to generate graphs with meta-communities for community detection algorithm benchmarking. The closest one is the LDBC-SNB data generator~\cite{Erling:2015:LSN:2723372.2742786} which has been provided by the generation of groups of users in the social network. Even though the generation process does not specifically enforce the generation of groups (meta-communities) for benchmarking community detection algorithms, the study from Prat-Pe\'rez and Dominguez-Sal reveals that these groups are more similar to the real meta-communities than those structural communities generated by the LFR benchmark~\cite{Prat-Perez:2014:CSS:2621934.2621942}.

The differences observed between structural and meta-communities reveal the need of more accurate community definitions tight more specifically to the domain or use case. Current community detection algorithms and graph generators for community detection are stuck to the traditional (and vague) definition of community, assuming that there exists a single algorithm that would fit all the use cases. Thus, future work requires the study of domain-specific community characteristics that can be used to generate graphs with a community structure that accurately resembles that of specific use cases, and thus revealing which are the best algorithms for each particular scenario.